### 2024-05-01 paper.qmd
Paper edits

---
title: "Estimating Frozen States in Civil Conflicts"
subtitle: "A Latent Variable Approach"
authors:
- name: J.L.A. Krusell
  email: "joshua.krusell@gu.se"
  affiliations:
      - "University of Gothenburg"
date: today
date-format: long
bibliography: library.bib
linestretch: 1.25
format:
    pdf:
      fontsize: 11pt
      papersize: A4
      include-in-header:
          - text: |
              \usepackage{graphicx}
              \usepackage{tikz}
      pdf-engine: lualatex
execute:
    cache: true
    echo: false
---

# Introduction

Civil conflicts are dynamics processes that may underfo several waves of escalation and de-escalation in intensity as the conflict oscillates between active and inactive phases of combat.

Identifying these periods is a crucial first step towards examining why certain conflicts become frozen

Civil conflict as a dynaic process with several wabes of escalation and de-escalation

  - How/why conflicts transition in and out of states of being frozen
    is still underexplored by researchers

"Why and when do some frozen conflicts re-escalate back into active conflicts?"

  - To be able to answer these questions we need to be able to
    identify patterns of escalation and de-escalation

In this article I propose a latent variable approach that treats the state of a conflict at a given time period as an unobserved variable

  - Evolves as a discrete markov process whose transition
    probabilities between frozen and active states is a time varying
    function of a set of covariates.

"By incorporating time varying covariates and hierarchical structures, the model captures the complex dynamics of conflict escalation and de-escalation, accounting for both short-term fluctuations and long-term trends."

This stands in direct contrast to the standard approach of coding frozen conflicts, and distinguishing between high and low intensity conflicts more broadly. Namely by identifying drops in battle related fatalities below pre-specified threshold.

  - Move away from arbitrary thresholds towards probabilistic
    inference of state labels that incorporate uncertainty.
  - At a cost of computational complexity

HMMs provide a probabilistic approach to capture the temporal shifts in conflict intensity and to estimate the underlying processes driving these transitions.

Better our understanding of both frozen conflicts, and conflict dynamics more generally

# Existing Measures of Frozen Conflicts

# HMMs in Conflict Research

# A Hidden Markov Model of Conflict Dynamics

## Model Specification

## Estimation

# Applied Example: Arms Transfers and Conflict Phases

## Data

Unit of analysis: Civil conflict sequences reaching at least 500 cumulative fatalities at the conflict-year level

Emission variable: battle-related fatalities

Independent variables:

- TIV (asinh)
- Ceasefire
- Peacekeeping operation
- Ongoing conflict
- Polyarchy (squared polynomial)
- Population (logged)
- GDPpc (logged)
- Duration (cubed polynomial)

## Priors/Identification

## Results

- Estimated state transitions w/ predicted emissions
- Beta parameter estimates for TIV
- Marginal effects:
  1. Probability of a frozen conflict **escalating**
  2. Probability of an active conflict **de-escalating**

# Conclusion

Shortcomings:
- Frozen conflict measure does not account for the ability of actors
  to re-escalate a conflict --- no distinction between a weakened
  insurgency versus a de facto separatist state (DOES NOT MEASURE REBEL CAPACITY)
- Excluded low intensity conflicts

Future work:
- incorporate more battlefield dynamics, including distance to pop. centers, proportion of calendar year with violent events etc etc.
- dyadic level rather than conflict level

# Acknowledgement

The computations were enabled by resources provided by the National Academic Infrastructure for Supercomputing in Sweden (NAISS), partially funded by the Swedish Research Council through grant agreement no. 2024/22-670.

\newpage

# References

::: {#refs}
:::

\newpage

# Appendix

1. Forward-backward algorithm (argue for why forward-backward instead of viterbi)
2. Simulation based calibration
3. ppc_dens_overlay plots
4. Table of \beta & \lambda estimates for all independent variables
5. Robustness checks